### 2022-02-25 outputs/paper/paper.Rmd
first commit

---
title: "Declining Birth rates in the US: An Analysis of Potential Factors"
subtitle: "Reproduction of 'The Puzzle of Falling US Birth Rates since the Great Recession' (Kearney, Levine & Pardue, 2022)"
author: Kimlin Chin
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Over the past 10 years, US birth rates have been dropping steadily to an all time low of 55.8 per 1,000 women in 2020. Kearney, Levine and Pardue (2022) analyzed this phenomenon by exploring demographic, economic, social and policy factors to little avail in enlightenment. However, they did discover a correlation between the decline and the birth cohorts of mothers, and offer the conjecture of shifting priorities as the reasoning. We replicate the results of this study with respect to the demographic and cohort effects, and discuss alternative interpretations of the data considering the impact of the feminist movement and the evolution of technology and media. The results were consistent with the initial study."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(haven)
library(kableExtra)
library(palmerpenguins)

# Read in data
fig1_data <- read_csv(here::here("inputs/data/fig_1.csv"))
fig2_data <- read_csv(here::here("inputs/data/figs_2a_2b.csv"))
birth_data <- full_join(fig1_data, fig2_data, by="year")

birth_cohort_data <- read_dta(here::here("inputs/data/nchs_cohort_analysis.dta"))
pop_cohort_data <- read_dta(here::here("inputs/data/agecomp-seer.dta"))
```

<!-- Possible papers: -->
<!-- - The Puzzle of Falling US Birth Rates since the Great Recession https://pubs.aeaweb.org/doi/pdfplus/10.1257/jep.36.1.151 -->
<!-- - Who Gets Their First Job at a Large Firm? The Distinct Roles of Education and Skills https://static1.squarespace.com/static/5c870f5194d71a8c5be53ae4/t/609c3108ef87b64561559084/1620848905329/pandp.20211009.pdf -->
<!-- - Family, Education, and Sources of Wealth among the Richest Americans, 1982-2012 https://www.aeaweb.org/articles?id=10.1257/aer.103.3.158&ArticleSearch%5Bwithin%5D%5Barticletitle%5D=1&ArticleSearch%5Bwithin%5D%5Barticleabstract%5D=1&ArticleSearch%5Bwithin%5D%5Bauthorlast%5D=1&ArticleSearch%5Bq%5D=education&JelClass%5Bvalue%5D=0&journal=&from=a&page=5&per-page=21&from=j -->

# Introduction

<!-- You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).  -->

<!-- Background / Broader Context -->
Birth rate is the number of live births per thousand in the population by year (Google dictionary). It is closely related to total fertility rate, which is the number of children that a hypothetical female would have over the course of her reproductive life [Statistics Canada](https://www23.statcan.gc.ca/imdb/p3Var.pl?Function=DEC&Id=45197). Birth rate is important since it is an indicator of population growth and determines the age structure of populations, which has various implications for the economy and society [@CLELAND2008364]. When the birth rate is too high or low, it affects all subgroups of the population in different ways, for example, a large number of new births will be a burden on the adult population to support them, and when this group ages to be elderly they will be a burden on governments to support them [@CLELAND2008364].

There are many possible factors that can affect birth rate. Recessions have been proposed to result in a drop in birth rates, specifically the Great Recession of 2007 [@pew; @focus; @econ]. @becker proposes an economic framework for fertility, where the "demand for children" is driven by the satisfaction that children brings to people weighted against factors such as preferences (e.g. religion, race, age), income, time, expenditure and means to support each child and other considerations (e.g. contraception, marriage). @og base their analysis of the declining US birth rates on these factors.

These analyses however, did not provide sufficient evidence to explain the decline in birth rates, with the authors stating "For any factor to have explained much of that decline, it would have had to change dramatically around the same time." [@og]. @og then went on to find a correlation between the recent birth cohorts of mothers and the substantial decline. In our reproduction of their paper, we build on their 'shifting priorities' explanation of the cohort effects by considering the other notable events that potentially may have influenced the nature of parenting of these cohorts, namely the Third wave of Feminism and the second half of the Digital Revolution.

We replicate the paper by @og with a focus on the following research questions:

* What is the trend in US birth rates over the period 1980-2020?
* How does the trend in birth rates vary with demographic factors such as age, race, education and marital status?
* How do the birth rates vary for different cohorts of mothers by their birth year?

While the original paper used Stata [@citeStata] for data processing and analysis, we use R [@citeR] for all data wrangling and analysis.

<!-- https://www.sciencedirect.com/science/article/pii/B9780123739605004731 -->
<!-- https://bmcpublichealth.biomedcentral.com/articles/10.1186/s12889-020-8331-7 -->

# Data

## Data Source and Methodology

The data on birth rates per 1,000 women ages 15-44 across all races and demographic population subgroups are obtained from the National Vital Statistics Reports for the years 2015, 2019 and 2020 [@2015birth; @2019birth; @2020birth], which they collected from birth certificates registered in all states and the District of Colombia. Aggregated data on the number of births for different cohorts of mothers by their age and birth year is provided by @og, calculated using public birth microdata across the period 1980-1989 from the NBER Natality Database and NCHS microdata from the period 1990-2019 (CDC NCHS 2020, NBER 2021b).

Single-age population counts, among all races from 1969-2019 and by race and Hispanic origin from 1990-2019, is obtained from the CDC SEER database (CDC NCI 2021).

## Attributes

The birth data consists of variables for the year, and birth rate for each population subgroup. For example, for race there are 3 variables: birth rate for whites, birth rate for blacks and birth rate for hispanics. A glimpse of the data can be seen in table \@ref(tab:birthdata).

```{r birthdata, echo=FALSE}
birth_data %>%
  filter(year > 1989) %>%
  slice(1:10) %>%
  kable(
    caption = "Ten rows of the birth rate data for years 1980-2020",
    col.names = c("Year", "All races", "Ages 15-19", "Ages 20-24", "Ages 25-29", "Ages 30-34", "Ages 35-39", "Ages 40-44", "White, non-Hispanic", "Black, non-Hispanic", "Hispanic"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )
```

For the aggregated cohort birth data, the variables are mother's age, birth cohort, number of births and cumulative births.

Here is a glimpse of the data:

```{r birthcohort, echo=FALSE}
birth_cohort_data %>%
  slice(1:10) %>%
  kable(
    caption = "First ten rows of the birth cohort data for 6 birth cohorts",
    col.names = c("Mother's age", "Cohort #", "Number of births", "Cumulative births"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )
```

For the population data, we process it.

```{r, echo=FALSE}
fig5_pop_data1 <- pop_cohort_data %>%
  select(1:32) %>%
  select(-c(stname)) %>%
  group_by(year) %>%
  summarise(across(everything(),sum)) # https://stackoverflow.com/questions/1660124/how-to-sum-a-variable-by-group

fig5_pop_data1 <- fig5_pop_data1 %>%
  pivot_longer(-c(year), names_to = "mage", values_to = "pop")

fig5_pop_data1 <- fig5_pop_data1 %>%
  mutate(mage = as.numeric(str_replace(mage, "fem", ""))) %>%
  mutate(cohort = year - mage) %>%
  mutate(cohort2 = case_when(cohort >= 1968 & cohort <=1972 ~ 1,
                             cohort >= 1973 & cohort <=1977 ~ 2,
                             cohort >= 1978 & cohort <=1982 ~ 3,
                             cohort >= 1983 & cohort <=1987 ~ 4,
                             cohort >= 1988 & cohort <=1992 ~ 5,
                             cohort >= 1993 & cohort <=1997 ~ 6))

fig5_pop_data1 <- na.omit(fig5_pop_data1)
fig5_pop_data1 <- fig5_pop_data1 %>%
  group_by(cohort2, mage) %>%
  summarise(pop = sum(pop))

```

```{r birthpop, echo=FALSE}
fig5_pop_data1 %>%
  filter(cohort2 == 1 & mage < 25) %>%
  kable(
    caption = "First ten rows of the population data for 6 birth cohorts",
    col.names = c("Cohort #", "Mother's age", "Population"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )
```

# Results

## Overall Trend

## Demographic Factors

## Cohorts

# Discussion

## Why is the declining birth rate relevant?

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

##

## Cohort effects - feminism, technological influence, media?

<!-- https://www.vox.com/2018/3/20/16955588/feminism-waves-explained-first-second-third-fourth -->
<!-- https://ourworldindata.org/fertility-rate -->
<!-- https://public.tableau.com/views/WorldDemographicsAnimated-LifeExpectancyvsFertilityRate/WorldDemographicsStoryline?:embed=y&:showVizHome=no&:display_count=y&:display_static_image=y&:bootstrapWhenNotified=true -->

<!-- https://www.nytimes.com/2021/06/27/opinion/falling-birthrate-women-babies.html -->
<!-- https://thefederalist.com/2018/10/02/low-birth-rates-parallel-not-just-economy-also-waves-feminism/ -->
<!-- https://unherd.com/2020/03/why-arent-women-having-more-babies/ -->

<!-- https://thenextweb.com/news/20-years-ago-today-the-world-wide-web-opened-to-the-public -->
<!-- https://www.forbes.com/sites/forbesbusinessdevelopmentcouncil/2020/10/14/recognizing-the-technological-revolution-and-preparing-for-the-next-economy/?sh=38724be83965 -->
<!-- https://en.wikipedia.org/wiki/Technological_revolution -->

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References